---
title: "act"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries.
```{r}
library(tidyverse)
library(infer)
```

***

You are trying to decide between two ACT prep classes. Class 1 begins with a general overview followed
by an extensive number of practice exams. Class 2 covers each topic in depth with practice questions 
arranged by topic. Sample data for each class is available in the file act.txt in your repository.  Based on a permutation test with
10,000 replications, does Class 1 increase ACT scores on average? Answer the same question for Class
2? Write a script named act.R or act.Rmd to address these questions. Underneath your code, write a 
conclusion as to which class (or, neither class) you would recommend. Justify your answers.

Read in data.
```{r}
act_data <- read.table('act.txt', header = TRUE)
```

Calculate difference in means for each class.
```{r}
(class_stats <- act_data %>%
   group_by(class) %>%
   summarize(avg_pre = mean(pre), 
             avg_post = mean(post),
             difference = avg_post - avg_pre))
```

Store true differences.
```{r}
c1_diff <- class_stats %>%
  filter(class == 'class1') %>%
  select(difference) %>%
  pull()

c2_diff <- class_stats %>%
  filter(class == 'class2') %>%
  select(difference) %>%
  pull()
```

Gather the 'pre' and 'post' columns in the data to make it tidy.
```{r}
act_data <- act_data %>%
  gather(key = score_period, value = score, -class)
```

Our null hypothesis will be that there is no difference in mean score before and after attending
each class. We can impose the null hypothesis by randomizing the post scores for each class and 
calculating the difference in means between pre and post scores. This will give us the distribution
of differences in means under the null hypothesis, which we can then compare our actual difference
in means to.

## Class 1

Create 'c1_data' by filtering for the obervations from the first class, and generating 10000 
replications of the original data. Then, create a new variable 'perm_score' by randomizing the 
scores. Group by each replication and the pre and post periods, then calculate the average pre and 
post class scores for each replication. Then calculate the difference between those average scores.
```{r}
c1_data <- act_data %>%
  filter(class == 'class1') %>%
  rep_sample_n(size = nrow(.),
               reps = 10000) %>%
  mutate(perm_score = sample(score)) %>%
  group_by(replicate, score_period) %>%
  summarize(avg_score = mean(perm_score)) %>%
  summarize(score_diff = diff(avg_score))
```

Graph the distribution of the difference in mean scores before and after the first class. Include
a verticle line showing the actual observed difference in mean score.
```{r}
c1_data %>%
  ggplot(aes(x = score_diff)) +
  geom_density() +
  geom_vline(xintercept = c1_diff)
```

Calculate the p-value by subtracting from 1 (since we're looking at the right tail of the 
distribution) the proportion of mean differences under the null that are less than our observed
difference.
```{r}
c1_data %>%
  summarize(p_value = 1 - sum(score_diff < c1_diff) / n())
```

We do not reject the null hypothesis that there is no difference in mean score before and after
taking the first class.

## Class 2

Create 'c2_data' by filtering for the obervations from the first class, and generating 10000 
replications of the original data. Then, create a new variable 'perm_score' by randomizing the 
scores. Group by each replication and the pre and post periods, then calculate the average pre and 
post class scores for each replication. Then calculate the difference between those average scores.
```{r}
c2_data <- act_data %>%
  filter(class == 'class2') %>%
  rep_sample_n(size = nrow(.),
               reps = 10000) %>%
  mutate(perm_score = sample(score)) %>%
  group_by(replicate, score_period) %>%
  summarize(avg_score = mean(perm_score)) %>%
  summarize(score_diff = diff(avg_score))
```

Graph the distribution of the difference in mean scores before and after the first class. Include
a verticle line showing the actual observed difference in mean score.
```{r}
c2_data %>%
  ggplot(aes(x = score_diff)) +
  geom_density() +
  geom_vline(xintercept = c2_diff)
```

Calculate the p-value by subtracting from 1 (since we're looking at the right tail of the 
distribution) the proportion of mean differences under the null that are less than our observed
difference.
```{r}
c2_data %>%
  summarize(p_value = 1 - sum(score_diff < c2_diff) / n())
```

We do not reject the null hypothesis that there is no difference in mean score before and after
taking the second class.

## Conclusion

Overall, I wouldn't recommend attending either class. The increase in mean score for both classes
is not statistically significant, so I would recommend spending your money on other study methods.
