---
title: "act"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries.
```{r}
library(tidyverse)
```

***

You are trying to decide between two ACT prep classes. Class 1 begins with a general overview 
followed by an extensive number of practice exams. Class 2 covers each topic in depth with practice 
questions arranged by topic. Sample data for each class is available in the file act.txt in your 
repository. Based on a permutation test with 10,000 replications, does Class 1 increase ACT scores on
average? Answer the same question for Class 2. Underneath your code, write a conclusion as to which 
class (or, neither class) you would recommend. Justify your answers.

***

For our permutation test, our null hypothesis will be that there is no difference in means for
our pre and post class scores.

Read in data.
```{r}
act_data <- read.table('act.txt', header = TRUE)
```

Calculate the average pre and post test scores for each class as well as the difference in average
pre and post scores for each class.
```{r}
(class_stats <- act_data %>%
   group_by(class) %>%
   summarize(avg_pre = mean(pre), 
             avg_post = mean(post),
             difference = avg_post - avg_pre))
```

Store true differences in average pre and post test scores for each class.
```{r}
c1_diff <- class_stats %>%
  filter(class == 'class1') %>%
  select(difference) %>%
  pull()

c2_diff <- class_stats %>%
  filter(class == 'class2') %>%
  select(difference) %>%
  pull()
```

# Function

Create a function, 'permuter' which filters the 'act_data' for the specified class. The 'nreps' 
arguement has a default value of 10,000, which is how many times we want to permute the data and
calculate our statistic of interest. After filtering for the specified class, the function generates
random draws between 0 and 1 from a uniform distribution. Whether those draws are above or below
0.5 determines how the values of 'pre' and 'post' test scores are permuted. Then, the function
calculates the average difference in test scores 'pre' and 'post' and returns that value.
```{r}
permuter <- function(which_class, nreps = 10000) {
  
  new_data <- act_data %>%
    filter(class == which_class)
  
  mean_diffs <- map_df(1:nreps, function(x) {
    new_data %>%
      mutate(x = runif(n()),
             perm_pre = ifelse(x > 0.5, pre, post),
             perm_post = ifelse(x < 0.5, pre, post)) %>%
      summarize(mean_diff = mean(perm_post - perm_pre))
      
  })
  
  return(mean_diffs)
  
}
```

# Class 1

Calculate 10,000 estimates of the difference in mean score under the null hypothesis for class one.
```{r}
c1_perm_diffs <- permuter('class1')
```

Graph the estimates and include a verticle line to show the difference for our class 1 sample.
```{r}
c1_perm_diffs %>%
  ggplot(aes(x = mean_diff)) +
  geom_density() +
  geom_vline(xintercept = c1_diff, color = 'red') +
  labs(title = 'Null Distribution of Mean Score Diff. - Class 1',
       x = 'Mean Score Difference',
       y = 'Density')
```

Calculate the p-value of our estimate for class one.
```{r}
c1_perm_diffs %>%
  summarize(P_Value = mean(mean_diff > c1_diff))
```

# Class 2

Calculate 10,000 estimates of the difference in mean score under the null hypothesis for class two.
```{r}
c2_perm_diffs <- permuter('class2')
```

Graph the estimates and include a verticle line to show the difference for our class 2 sample.
```{r}
c2_perm_diffs %>%
  ggplot(aes(x = mean_diff)) +
  geom_density() +
  geom_vline(xintercept = c2_diff, color = 'red') +
  labs(title = 'Null Distribution of Mean Score Diff. - Class 2',
       x = 'Mean Score Difference',
       y = 'Density')
```

Calculate the p-value of our estimate for class two.
```{r}
c2_perm_diffs %>%
  summarize(P_Value = mean(mean_diff > c2_diff))
```

Based on this analysis, I would recommend that someone take class two. The increase in average score
that it produces is significant at the 5% level. However, our point estimate of the increase in 
average score for class 1 is larger than the point estimate for class 2. We only have 20 data points
from class 1, so if we were to get more, we might be able to show that the increase in score that
class 1 produces is statistically greater than zero. Until that point, I would recommend class 2.
